---
title: "PitchRx Code"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warnings=FALSE}
#####################FOURSEAM############
library(dplyr)
Fourseam_FB<-read.csv("4F_1.csv",  stringsAsFactors=FALSE, header=T)
seq<-c("4F_2.csv","4F_3.csv","4F_4.csv","4F_5.csv","4F_6.csv","4F_7.csv","4F_8.csv","4F_9.csv")
for (val in seq){
  new<-read.csv(val, stringsAsFactors=FALSE, header=T)
  Fourseam_FB<-rbind(Fourseam_FB, new)
}


Twoseam_FB<-read.csv("2F_1.csv",  stringsAsFactors=FALSE, header=T)
seq<-c("2F_2.csv","2F_3.csv","2F_4.csv","2F_5.csv","2F_6.csv","2F_7.csv","2F_8.csv","2F_9.csv")
for (val in seq){
  new<-read.csv(val, stringsAsFactors=FALSE, header=T)
  Twoseam_FB<-rbind(Twoseam_FB, new)
}


Change<-read.csv("CH_1.csv",  stringsAsFactors=FALSE, header=T)
seq<-c("CH_2.csv","CH_3.csv","CH_4.csv","CH_5.csv","CH_6.csv","CH_7.csv","CH_8.csv","CH_9.csv")
for (val in seq){
  new<-read.csv(val, stringsAsFactors=FALSE, header=T)
  Change<-rbind(Change, new)
}

Sinker<-read.csv("SN_1.csv",  stringsAsFactors=FALSE, header=T)
seq<-c("SN_2.csv","SN_3.csv","SN_4.csv","SN_5.csv","SN_6.csv","SN_7.csv","SN_8.csv","SN_9.csv")
for (val in seq){
  new<-read.csv(val, stringsAsFactors=FALSE, header=T)
  Sinker<-rbind(Sinker, new)
}

Slider<-read.csv("SL_1.csv",  stringsAsFactors=FALSE, header=T)
seq<-c("SL_2.csv","SL_3.csv","SL_4.csv","SL_5.csv","SL_6.csv","SL_7.csv","SL_8.csv","SL_9.csv")
for (val in seq){
  new<-read.csv(val, stringsAsFactors=FALSE, header=T)
  Slider<-rbind(Slider, new)
}

KCurve<-read.csv("KC_1.csv",  stringsAsFactors=FALSE, header=T)
seq<-c("KC_2.csv","KC_3.csv","KC_4.csv","KC_5.csv","KC_6.csv","KC_7.csv","KC_8.csv","KC_9.csv")
for (val in seq){
  new<-read.csv(val, stringsAsFactors=FALSE, header=T)
  KCurve<-rbind(KCurve, new)
}

Splitter<-read.csv("FS_1.csv",  stringsAsFactors=FALSE, header=T)
seq<-c("FS_2.csv","FS_3.csv","FS_4.csv","FS_5.csv","FS_6.csv","FS_7.csv","FS_8.csv","FS_9.csv")
for (val in seq){
  new<-read.csv(val, stringsAsFactors=FALSE, header=T)
  Splitter<-rbind(Splitter, new)
}

Cutter<-read.csv("FC_1.csv",  stringsAsFactors=FALSE, header=T)
seq<-c("FC_2.csv","FC_3.csv","FC_4.csv","FC_5.csv","FC_6.csv","FC_7.csv","FC_8.csv","FC_9.csv")
for (val in seq){
  new<-read.csv(val, stringsAsFactors=FALSE, header=T)
  Cutter<-rbind(Cutter, new)
}

Curve<-read.csv("CV_1.csv",  stringsAsFactors=FALSE, header=T)
seq<-c("CV_2.csv","CV_3.csv","CV_4.csv","CV_5.csv","CV_6.csv","CV_7.csv","CV_8.csv","CV_9.csv")
for (val in seq){
  new<-read.csv(val, stringsAsFactors=FALSE, header=T)
  Curve<-rbind(Curve, new)
}

Pitches<-rbind(Fourseam_FB, Twoseam_FB, Change, Curve, KCurve, Slider, Cutter, Splitter, Sinker)


library(ggplot2)
library(dplyr)
library(stringr)
library(knitr)
library(tidyr)
library(shiny)
library(DT)
library(caret)
library(plotly)
Pitches_Wrangled<-Pitches%>%
  select(pfx_x, pfx_z, effective_speed,release_speed ,release_spin_rate, pitch_type, p_throws,player_name, description, stand, woba_value, estimated_woba_using_speedangle, launch_speed, launch_angle)%>%
  mutate(pfx_x=as.numeric(pfx_x), pfx_z=as.numeric(pfx_z), effective_speed=as.numeric(effective_speed),release_speed=as.numeric(release_speed),release_spin_rate=as.numeric(release_spin_rate), woba_value=as.numeric(woba_value), estimated_woba_using_speedangle=as.numeric(estimated_woba_using_speedangle), launch_speed=as.numeric(launch_speed), launch_angle=as.numeric(launch_angle))%>%
  drop_na(pfx_x)%>%
  drop_na(pfx_z)%>%
  drop_na(effective_speed)%>%
  drop_na(release_speed)%>%
  drop_na(release_spin_rate)%>%
  filter(description!="pitchout")
  mutate(description=as.factor(description))





```



```{r}
library(ggplot2)
library(dplyr)
library(stringr)
library(knitr)
library(tidyr)
library(shiny)
library(DT)
library(caret)

#UI Side
ui<-shinyUI(navbarPage("PitchFX APP",
                       tabPanel("App Explanation"),
                       tabPanel("Isaac", titlePanel("User Input"),
                                sidebarLayout(
                                  sidebarPanel(helpText("PitchFX Input"),
                                               selectInput(inputId="pitcher_throws",
                                                           label="Pitcher Handedness",
                                                           choices=list("Right-Handed"="R",
                                                                        "Left-Handed"="L")),
                                               textInput(inputId="pfx_x",
                                                         label="PFX_X",
                                                         value=0),
                                               textInput(inputId="pfx_z",
                                                         label="PFX_Z",
                                                         value=0),
                                               textInput(inputId="effective_speed",
                                                         label="Effective Velo",
                                                         value=90),
                                               textInput(inputId="release_speed",
                                                         label="Release Velo",
                                                         value=90),
                                               textInput(inputId="release_spin_rate",
                                                         label="Release Spin Rate",
                                                         value=2000),
                                               sliderInput(inputId="pick_md", 
                                                           label="Pick Mahalanobis Distance",
                                                           min = 0, max = 15,
                                                           value = 1.0, step = 0.1),
                                               checkboxGroupInput(inputId = "batter_hand",
                                                                  label="Batter Handedness",
                                                                  choices=list("Right-Handed"="R",
                                                                               "Left-Handed"="L"),
                                                                  selected=c("R", "L")),
                                               actionButton("go", "Fit Models")
                                               
                                  ), 
                                  mainPanel(
                                    tabsetPanel(type="tabs",
                                                tabPanel("One", DT::dataTableOutput("pitchTable")),
                                                tabPanel("Two",DT::dataTableOutput("swingTable")),
                                                tabPanel("Three", DT::dataTableOutput("models")),
                                                tabPanel("Scatter Plot", plotlyOutput("plot1")),
                                                tabPanel("Swing and Miss", plotlyOutput("plot2")),
                                                 tabPanel("Hist of wOBA", plotlyOutput("plot3")),
                                                tabPanel("dt", dataTableOutput("pitchTable1"))
                                                )
                                  )
                                ))
                                
                       )
                       
)
#server
server<-shinyServer(function(input,output){

    pitch_data_for_cov<-reactive({
      Pitches_Wrangled[,1:5]
    })
    
    cov_matrix<-reactive({
      
      solve(cov(pitch_data_for_cov(), y=NULL, use="everything"))
    })
    
    pitch_data<-reactive({
      CovInv<-cov_matrix()
      Pitches_Wrangled%>%
        mutate(d1=pfx_x-as.numeric(input$pfx_x))%>%
        mutate(d2=pfx_z-as.numeric(input$pfx_z))%>%
        mutate(d3=effective_speed-as.numeric(input$effective_speed))%>%
        mutate(d4=release_speed-as.numeric(input$release_speed))%>%
        mutate(d5=release_spin_rate-as.numeric(input$release_spin_rate))%>%
        mutate(p1=d1*CovInv[1]+d2*CovInv[2]+d3*CovInv[3]+d4*CovInv[4]+d5*CovInv[5])%>%
        mutate(p2=d1*CovInv[6]+d2*CovInv[7]+d3*CovInv[8]+d4*CovInv[9]+d5*CovInv[10])%>%
        mutate(p3=d1*CovInv[11]+d2*CovInv[12]+d3*CovInv[13]+d4*CovInv[14]+d5*CovInv[15])%>%
        mutate(p4=d1*CovInv[16]+d2*CovInv[17]+d3*CovInv[18]+d4*CovInv[19]+d5*CovInv[20])%>%
        mutate(p5=d1*CovInv[21]+d2*CovInv[22]+d3*CovInv[23]+d4*CovInv[24]+d5*CovInv[25])%>%
        mutate(md2=(d1*p1+d2*p2+d3*p3+d4*p4+d5*p5))%>%
        mutate(md=(md2)^.5)%>%
        arrange(md)%>%
        filter(md<=input$pick_md)%>%
        filter(stand==input$batter_hand)%>%
        filter(p_throws==input$pitcher_throws)
      
    })
    swing_data<-reactive({
      pitch_data()%>%
        filter(description=="foul"|description=="foul_tip"|description=="hit_into_play"|description=="hit_into_play_no_out"|description=="hit_into_play_score"|description=="swinging_strike"|description=="swinging_strike_blocked")%>%
        mutate(estimated_woba_using_speedangle=ifelse(is.na(estimated_woba_using_speedangle),0,estimated_woba_using_speedangle))
    })
    
    
    
    output$pitchTable<-DT::renderDataTable({
      DT::datatable(data=pitch_data(), rownames=FALSE)
    }, server=TRUE)
    
    

    output$swingTable<-DT::renderDataTable({
      DT::datatable(data=swing_data(), rownames=FALSE)
    }, server=TRUE)
    
    
    
    ###############
    input_dataframe<-reactive({
      data.frame("pfx_x"=input$pfx_x, "pfx_z"=input$pfx_z, "effective_speed"=input$effective_speed, "release_speed"=input$release_speed, "release_spin_rate"=input$release_spin_rate)
    })
    
    
    
    
    
    
    
    
    ############
    
    
    
    ####################################CHARLIE YOUR SERVER CODE HERE#############################################
    library(plotly)
    pitch_data_for_scatter<-reactive({
      pitch_data() %>%
        filter(description != "missed_bunt") %>%
        filter(description != "ball") %>%
        filter(description != "swinging_strike") %>%
        filter(description != "called_strike") %>%
        filter(description != "foul_tip") %>%
        filter(description != "foul_bunt") %>%
        filter(description != "foul") %>%
        filter(description != "hit_by_pitch") %>%
        filter(description != "swinging_strike_blocked") %>%
       filter(description != "blocked_ball")
      }) 
    
    pitch_data_for_SwingAndMiss<-reactive({
      pitch_data() %>%
        filter(description != "missed_bunt") %>%
        filter(description != "foul_bunt") %>%
        mutate(outcome = ifelse(description == "foul_tip" | description == "foul" | description == "hit_into_play" | description == "hit_into_play_score" | description == "hit_into_play_no_out", "Contact", ifelse(description == "swinging_strike" | description == "swinging_strike_blocked", "Swing_and_Miss", ifelse(description == "ball" | description == "called_strike" | description == "blocked_ball" | description == "hit_by_pitch", "Taken_Pitch", 0))))
                })
      
    output$pitchTable1<-DT::renderDataTable({
      DT::datatable(data=pitch_data_for_SwingAndMiss(), rownames=FALSE)
    }, server=TRUE)
    
    output$plot1 <- renderPlotly({
      req(input$pitcher_throws, input$pfx_x, input$pfx_z, input$effective_speed, input$release_spin_rate, input$release_speed, input$pick_md, input$batter_hand)
      ggplot(data = pitch_data_for_scatter(), 
           mapping = aes_string(x = "launch_angle", y = "launch_speed")) + geom_point(col = "red")+
        ylab("Launch Speed") + xlab("Launch Angle") + labs(title = "Launch Angle versus Launch Speed")
    })
    
    output$plot2 <- renderPlotly({
      req(input$pitcher_throws, input$pfx_x, input$pfx_z, input$effective_speed, input$release_spin_rate, input$release_speed, input$pick_md, input$batter_hand)
      ggplot(data = pitch_data_for_SwingAndMiss(), 
           mapping = aes(outcome)) + geom_bar(fill = "red")+
        ylab("Amount") + xlab("Outcome") + labs(title = "Chances of Making Contact")
    })
    
    output$plot3 <- renderPlotly({
      req(input$pitcher_throws, input$pfx_x, input$pfx_z, input$effective_speed, input$release_spin_rate, input$release_speed, input$pick_md, input$batter_hand)
      ggplot(data = pitch_data_for_SwingAndMiss(), 
           mapping = aes(woba_value)) + geom_bar(fill = "red")+
        ylab("Amount") + xlab("wOBA Value") + labs(title = "Histogram of wOBA Values")
    })
    
    #########################################################################################################
})




shinyApp(ui=ui, server=server)

```




    cv_opts<-trainControl(method="cv", number=10)
    
    
    






```{r}

Fourseam_FB<-read.csv("4F_1.csv",  stringsAsFactors=FALSE, header=T)
seq<-c("4F_2.csv","4F_3.csv","4F_4.csv","4F_5.csv","4F_6.csv","4F_7.csv","4F_8.csv","4F_9.csv")
for (val in seq){
  new<-read.csv(val, stringsAsFactors=FALSE, header=T)
  Fourseam_FB<-rbind(Fourseam_FB, new)
}

FourSeam_Compare<-Fourseam_FB%>%
  select(pfx_x, pfx_z, effective_speed,release_speed ,release_spin_rate, player_name, description, woba_value)%>%
  mutate(pfx_x=as.numeric(pfx_x), pfx_z=as.numeric(pfx_z), effective_speed=as.numeric(effective_speed),release_speed=as.numeric(release_speed),release_spin_rate=as.numeric(release_spin_rate), woba_value=as.numeric(woba_value))%>%
  drop_na(pfx_x)%>%
  drop_na(pfx_z)%>%
  drop_na(effective_speed)%>%
  drop_na(release_speed)%>%
  drop_na(release_spin_rate)

pitch_choices<-(Fourseam_FB_Wrangled)

FourSeam_Cov<-cov(FourSeam_Compare[,1:5], y=NULL, use="everything")
CovInv<-solve(FourSeam_Cov)
x<-c(0, 0, 90, 93, 2000)
x2<-c(-1.3185, 1.0058, 93.948, 96.0, 2192)
ll<-I(list(c(-.7521, 1.412, 91.631, 96.2, 2012)))

mahalanobis(as.numeric(x), as.numeric(FourSeam_Compare[2,c(1,2,3,4,5)]), FourSeam_Cov)
mahalanobis(as.numeric(ll[[1]]), as.numeric(FourSeam_Compare[1,c(1,2,3,4,5)]), FourSeam_Cov)



list <- as.vector(split(FourSeam_Compare[1:20,1:5], seq(nrow(FourSeam_Compare[1:20,1:5]))))
list2<-unlist(list, use.names = FALSE)

FourSeam_Compare2<-FourSeam_Compare%>%
  #mutate(subvec=I(list(x)[]))%>%
  #mutate(lists=list[])%>%
  #mutate(lists=unlist(lists[], use.names=FALSE))%>%
  #mutate(fx=unlist(lists))
 # mutate(md=mahalanobis(as.numeric(c(subvec[[1]][1], subvec[[1]][2], subvec[[1]][3], subvec[[1]][4], subvec[[1]][5])),
                        #as.numeric(c(lists[[1]][1], 1, 91, 92, 2000)),
                        #FourSeam_Cov))%>%
  mutate(d1=pfx_x-x[1])%>%
  mutate(d2=pfx_z-x[2])%>%
  mutate(d3=effective_speed-x[3])%>%
  mutate(d4=release_speed-x[4])%>%
  mutate(d5=release_spin_rate-x[5])%>%
  mutate(p1=d1*CovInv[1]+d2*CovInv[2]+d3*CovInv[3]+d4*CovInv[4]+d5*CovInv[5])%>%
  mutate(p2=d1*CovInv[6]+d2*CovInv[7]+d3*CovInv[8]+d4*CovInv[9]+d5*CovInv[10])%>%
  mutate(p3=d1*CovInv[11]+d2*CovInv[12]+d3*CovInv[13]+d4*CovInv[14]+d5*CovInv[15])%>%
  mutate(p4=d1*CovInv[16]+d2*CovInv[17]+d3*CovInv[18]+d4*CovInv[19]+d5*CovInv[20])%>%
  mutate(p5=d1*CovInv[21]+d2*CovInv[22]+d3*CovInv[23]+d4*CovInv[24]+d5*CovInv[25])%>%
  mutate(md2=(d1*p1+d2*p2+d3*p3+d4*p4+d5*p5))%>%
  mutate(md=(md2)^.5)

  
  
FourSeam_Compare$md<-0
  
for (ii in 255801){
  FourSeam_Compare$md[ii]<-mahalanobis(as.numeric(x), as.numeric(FourSeam_Compare[1,c(1,2,3,4,5)]), FourSeam_Cov)
}


```


```{r}
Twoseam_FB<-read.csv("2F_1.csv",  stringsAsFactors=FALSE, header=T)
seq2<-c("2F_2.csv","2F_3.csv","2F_4.csv","2F_5.csv","2F_6.csv","2F_7.csv","2F_8.csv","2F_9.csv")
for (val in seq2){
  new<-read.csv(val, stringsAsFactors=FALSE, header=T)
  Twoseam_FB<-rbind(Twoseam_FB, new)
}


Test2<-Twoseam_FB%>%
  for (ii in 1:3){
    mutate(paste("il", ii), 1)
  }


```